---
title: "FishMIP_validation"
author: "Phoebe.Woodworth-Jefcoats@noaa.gov"
date: '2022-11-10'
output: html_document
---

## Compare catch at size

This validation compares catch in numbers across body sizes.  This is pretty much the only validation data available at this level of detail for this fishery.

Set up the environment:
```{r}
# Load libraries
library("remotes")
remotes::install_github("sizespectrum/therMizer")
library("therMizer")
```

Set up the species parameters, just to make some things easier:
```{r}
# Load parameters 
HIparams <- read.csv("HIregion_species_params.csv", header = TRUE)
HIinter <- read.csv("HIregion_inter.csv", row.names = 1)
HIinter <- as(HIinter,"matrix")
HIgears <- read.csv("HIregion_gears_params.csv", header = TRUE)

# Generate parameters
params <- newMultispeciesParams(HIparams, interaction = HIinter, kappa = 1e12, min_w_pp = 1e-14, w_pp_cutoff = 455400) 
```


Load the validation simulation:
```{r}
load("sim_steady_validate.Rdata")
```

Load observed data that will be used to validate the model:
```{r}
# Load observed catch-at-size
obsDirectory <- "/Users/phoebe.woodworth/Documents/Integrated Model/HawaiiLongline"
albacore <- read.table(paste(obsDirectory,"/albacore_N.dat", sep = ""), header = FALSE)
bigeye <- read.table(paste(obsDirectory,"/bigeye_N.dat", sep = ""), header = FALSE)
blue_marlin <- read.table(paste(obsDirectory,"/blue_marlin_N.dat", sep = ""), header = FALSE)
blue_shark <- read.table(paste(obsDirectory,"/blue_shark_N.dat", sep = ""), header = FALSE)
lancetfish <- read.table(paste(obsDirectory,"/lancetfish_N.dat", sep = ""), header = FALSE)
mahi <- read.table(paste(obsDirectory,"/mahi_N.dat", sep = ""), header = FALSE)
opah <- read.table(paste(obsDirectory,"/opah_N.dat", sep = ""), header = FALSE)
skipjack <- read.table(paste(obsDirectory,"/skipjack_N.dat", sep = ""), header = FALSE)
striped_marlin <- read.table(paste(obsDirectory,"/striped_marlin_N.dat", sep = ""), header = FALSE)
swordfish <- read.table(paste(obsDirectory,"/swordfish_N.dat", sep = ""), header = FALSE)
wahoo <- read.table(paste(obsDirectory,"/wahoo_N.dat", sep = ""), header = FALSE)
yellowfin <- read.table(paste(obsDirectory,"/yellowfin_N.dat", sep = ""), header = FALSE)

# Aggregate species to speed things up
observedN <- array(numeric(),c(12,100)) # species x size
observedN[1,] <- bigeye[,2]
observedN[2,] <- mahi[,2]
observedN[3,] <- blue_shark[,2]
observedN[4,] <- skipjack[,2]
observedN[5,] <- yellowfin[,2]
observedN[6,] <- albacore[,2]
observedN[7,] <- striped_marlin[,2]
observedN[8,] <- wahoo[,2]
observedN[9,] <- swordfish[,2]
observedN[10,] <- blue_marlin[,2]
observedN[11,] <- lancetfish[,2]
observedN[12,] <- opah[,2]
```

Determine yield-at-size:
```{r}
# Write a function to get yield by number at size
# This borrowed from other mizer functions
getYieldSize <- function(object) {
  no_sp <- nrow(params@species_params)
  min_w <- rep(min(params@w),no_sp)
  max_w <- rep(max(params@w),no_sp)
  
  min_n <- plyr::aaply(min_w, 1, function(x) params@w >= x, .drop = FALSE)
  max_n <- plyr::aaply(max_w, 1, function(x) params@w <= x, .drop = FALSE)
  size_range <- min_n & max_n
  dimnames(size_range) <- list(sp = params@species_params$species, 
                             w = signif(params@w, 3)) 
  
	if (is(object, "MizerSim")) {
		sim <- object
		n_at_size <- sweep(sim@n, c(2,3), size_range, '*')
		f_gear <- getFMortGear(sim)
		f_gear_sz <- sweep(f_gear, c(3,4),size_range,'*')
		yield_gear_species_size <- sweep(f_gear_sz, c(1,3,4), n_at_size,'*')
		return(apply(yield_gear_species_size,c(1,3,4),sum))
	}
	stop("'object' should be a MizerSim object")
}

# Yield at size from simulation
yield_at_size <- getYieldSize(sim_steady_validate)
```

Compare modeled and observed catch, numerically and visually:
```{r}
# Create arrays for r and p-value
rValues <- array(numeric(),c(12)) 
pValues <- array(numeric(),c(12)) 

for (sp in seq(1,12,1)) {
		calib_test <- cor.test(observedN[sp,], yield_at_size[40,sp,])
		
		rValues[sp] <- calib_test$estimate
		
		pValues[sp] <- calib_test$p.value

	}

# Helpful things for plotting
species <- species_params(params)$species

# Plots
par(mfrow = c(3,4))
for (sp in seq(1,12,1)) {
	plot(log10(params@w), log10(yield_at_size[40,sp,]), type = "l", col = "darkgreen", xlim = c(-2,6), ylim = c(-3,10))
	par(new = TRUE)
	
	plot(log10(params@w), log10(observedN[sp,]), type = "l", col = "black", xlim = c(-2,6), ylim = c(-3,10))
	par(new = TRUE)
	plot(log10(params@w), log10(observedN[sp,]/0.06), type = "l", lty = 2,col = "black", xlim = c(-2,6), ylim = c(-3,10), main = species[sp])
	text(0, 9, paste("r = ",signif(rValues[sp],3)))
	text(0, 7, paste("p = ",signif(pValues[sp],3)))
	# dev.off()
}

# The observed catch is roughly 20% of the total catch.
# And every third fish is measured.
# Hence the '/0.06' for the plot (0.2 * 0.3 = 0.06)
# The observed lengths were converted to weights.
# See Woodworth-Jefcoats et al. 2019 for l-w conversion references.
```
All correlations are significant (max p-value = 0.00278638).  r-values range from 0.2960262 to 0.922207, with a mean of 0.6505192 and a median of 0.684653.  The comparison to observed catch looks about as good as it has for previous iterations.  Some species are better than others, as has been the case.  Improved understanding of species' size structure, life history, and stock-recruit relationships could possibly improve model fit.  Also, note that the data from this portion of the calibration are form 2006 - 2015, _after_ the permitted window of years for calibration.  Data are not available prior to 2006.