---
title: "package_output"
author: "Phoebe.Woodworth-Jefcoats@noaa.gov"
date: '2022-11-10'
output: html_document
---

## Package output

This script is used to package up the therMizer output for submission to ISIMIP.

Set up the environment:
```{r}
# Load libraries
library("remotes")
remotes::install_github("sizespectrum/therMizer")
library("therMizer")
library("lubridate") # for the time identifiers
library("tidyverse") # for combining numbers and dates, and for writing files
```

Load needed information:
```{r}
# These are the species-specific weights at 30 and 90 cm.
wgts_30cm <- read.table("wgts_30cm.dat")
wgts_90cm <- read.table("wgts_90cm.dat")

# This is the domain area, in square meters:
area <- 1.946 * 10^13

# Set up time steps that reflect actual months used, 
# noting that ISIMIP requests time as months since 1 January 1840
t_start <- ((1961 - 1840) * 12) - 1
time_steps <- seq(t_start, t_start + 599, 1)

# And time in year-month-day format
ymd_start <- ymd(19610101)
ymd_steps <- c(ymd_start)
t <- 1
while (t < 600) {
  ymd_next <- ymd_steps[length(ymd_steps)] + months(1)
  ymd_steps <- c(ymd_steps, ymd_next)
  t <- t + 1
}
```

Load model output:
```{r}
load("projection_steady_F.Rdata")
load("projection_steady_noF.Rdata")
```

Some helpful reference points:
```{r}
min_w <- min(projection_steady_F@params@w)
max_w <- max(projection_steady_F@params@w)
sp_w_inf <- projection_steady_F@params@species_params$w_inf
```

Write some helpful functions:
```{r}
# Taking advantage of mizer's built in function for this
# Writing functions that divide biomass and catch by area and sums across all species
# Note that the timesteps of interest (1:600, omitting time zero) are coded into the functions
# This saved me time later, but may not be the best approach.
# These functions also require a list of numeric species indices.

# Biomass - total and for size bins
tcb <- function(scenario, min_wt, max_wt, sp){
	consumer_biomass <- getBiomass(scenario, min_w = min_wt, max_w = max_wt)
	
	if (length(sp) > 1) {
	  consumer_gm2 <- apply(sweep(consumer_biomass[2:601,sp], 2, area, "/"), 1, sum)
	} else {
	  consumer_gm2 <- consumer_biomass[2:601,sp] / area
	}
	
	return(consumer_gm2)
}

# Catch - total
tc <- function(scenario, sp){
  catch <- getYield(scenario)
  
  if (length(sp) > 1) {
    catch_gm2 <- apply(sweep(catch[2:601,sp], 2, area, "/"), 1, sum)
  } else {
    catch_gm2 <- catch[2:601,sp] / area
  }
  
  return(catch_gm2)
}

# Catch in size bins
tc_size_bins <- function(scenario, min_wt, max_wt) {
  
  # Sizes of interest
  sz <- which(scenario@params@w >= min_wt & scenario@params@w <= max_wt)
 
  # Fishing mortality
  fishing_mortality <- getFMort(scenario)
  
  # Catch in numbers at size
  catch_n <- sweep(scenario@n[2:601,,sz], c(1,2,3), fishing_mortality[2:601,,sz], "*")
  
  # Biomass of catch at size summed across all sizes
  catch_sp <- apply(sweep(catch_n, 3, scenario@params@w[sz] * scenario@params@dw[sz], "*"), c(1,2), sum)
  
  # Sum across all species and divide by area
  catch_gm2 <- apply(sweep(catch_sp, 2, area, "/"), 1, sum)
  
  return(catch_gm2)
}

```

Create requested variables, plot them to make sure they're reasonable, and save them.

Total Consumer & Pelagic Biomass Density (tcb & tpb), all consumers:
```{r}
# Note: the 1:12 is to include all 12 species

tcb_histsoc <- tibble(time_steps, ymd_steps, tcb(projection_steady_F, min_w, max_w, 1:12))
colnames(tcb_histsoc) <- c("Months.Since.1Jan1840", "Year.Month.Day", "tcb.gm-2")
tcb_nat <- tibble(time_steps, ymd_steps, tcb(projection_steady_noF, min_w, max_w, 1:12))
colnames(tcb_nat) <- c("Months.Since.1Jan1840", "Year.Month.Day", "tcb.gm-2")

tpb_histsoc <- tibble(time_steps, ymd_steps, tcb(projection_steady_F, min_w, max_w, 1:12))
colnames(tpb_histsoc) <- c("Months.Since.1Jan1840", "Year.Month.Day", "tpb.gm-2")
tpb_nat <- tibble(time_steps, ymd_steps, tcb(projection_steady_noF, min_w, max_w, 1:12))
colnames(tpb_nat) <- c("Months.Since.1Jan1840", "Year.Month.Day", "tpb.gm-2")
```

Total consumer biomass density in log10 weight bins (tcblog10):
```{r}
tcblog10_histsoc <- tibble(time_steps,
                           ymd_steps,
                           tcb(projection_steady_F, 10^0, 10^1, 1:12),
                           tcb(projection_steady_F, 10^1, 10^2, 1:12),
                           tcb(projection_steady_F, 10^2, 10^3, 1:12),
                           tcb(projection_steady_F, 10^3, 10^4, 1:12),
                           tcb(projection_steady_F, 10^4, 10^5, 1:12), 
                           tcb(projection_steady_F, 10^5, max_w, 1:12))
colnames(tcblog10_histsoc) <- c("Months.Since.1Jan1840", 
                                "Year.Month.Day", 
                                "1to10g.gm-2",
                                "10to100g.gm-2",
                                "100gto1kg.gm-2",
                                "1to10kg.gm-2",
                                "10to100kg.gm-2",
                                "GT100kg.gm-2")

tcblog10_nat <- tibble(time_steps,
                           ymd_steps,
                           tcb(projection_steady_noF, 10^0, 10^1, 1:12),
                           tcb(projection_steady_noF, 10^1, 10^2, 1:12),
                           tcb(projection_steady_noF, 10^2, 10^3, 1:12),
                           tcb(projection_steady_noF, 10^3, 10^4, 1:12),
                           tcb(projection_steady_noF, 10^4, 10^5, 1:12), 
                           tcb(projection_steady_noF, 10^5, max_w, 1:12))
colnames(tcblog10_nat) <- c("Months.Since.1Jan1840", 
                                "Year.Month.Day", 
                                "1to10g.gm-2",
                                "10to100g.gm-2",
                                "100gto1kg.gm-2",
                                "1to10kg.gm-2",
                                "10to100kg.gm-2",
                                "GT100kg.gm-2")
```

Total catch density and total industrial catch density (tc & tic):
```{r}
tc_histsoc <- tibble(time_steps, ymd_steps, tc(projection_steady_F, 1:12))
colnames(tc_histsoc) <- c("Months.Since.1Jan1840", "Year.Month.Day", "tc.gm-2")

tic_histsoc <- tibble(time_steps, ymd_steps, tc(projection_steady_F, 1:12))
colnames(tic_histsoc) <- c("Months.Since.1Jan1840", "Year.Month.Day", "tic.gm-2")
```

Total catch density in log10 weight bins (tclog10):
```{r}
tclog10_histsoc <- tibble(time_steps,
                           ymd_steps,
                           tc_size_bins(projection_steady_F, 10^0, 10^1),
                           tc_size_bins(projection_steady_F, 10^1, 10^2),
                           tc_size_bins(projection_steady_F, 10^2, 10^3),
                           tc_size_bins(projection_steady_F, 10^3, 10^4),
                           tc_size_bins(projection_steady_F, 10^4, 10^5), 
                           tc_size_bins(projection_steady_F, 10^5, max_w))
colnames(tclog10_histsoc) <- c("Months.Since.1Jan1840", 
                                "Year.Month.Day", 
                                "1to10g.gm-2",
                                "10to100g.gm-2",
                                "100gto1kg.gm-2",
                                "1to10kg.gm-2",
                                "10to100kg.gm-2",
                                "GT100kg.gm-2")
```

Optional output: biomass and catch density of small pelagics with L infinity <30cm (bp30cm & cp30cm):
```{r}
# This model has no species that fit this category
sp_LT30cm <- which(wgts_30cm[,2] > sp_w_inf)
```

Optional output: biomass and catch density of medium pelagics with 30cm<=L_infinity<90cm (bp30to90cm & cp30to90cm):
```{r}
# Identify species with L infinity < 90cm
sp3090 <- which(wgts_90cm[,2] > sp_w_inf)

bp30to90cm_histsoc <- tibble(time_steps, ymd_steps, tcb(projection_steady_F, min_w, max_w, sp3090))
colnames(bp30to90cm_histsoc) <- c("Months.Since.1Jan1840", "Year.Month.Day", "bp30to90cm.gm-2")
bp30to90cm_nat <- tibble(time_steps, ymd_steps, tcb(projection_steady_noF, min_w, max_w, sp3090))
colnames(bp30to90cm_nat) <- c("Months.Since.1Jan1840", "Year.Month.Day", "bp30to90cm.gm-2")

cp30to90cm_histsoc <- tibble(time_steps, ymd_steps, tc(projection_steady_F, sp3090))
colnames(cp30to90cm_histsoc) <- c("Months.Since.1Jan1840", "Year.Month.Day", "cp30to90cm.gm-2")
```

Optional output: biomass and catch density of large pelagics with L infinity >=90 cm (bp90cm & cp90cm)
```{r}
sp90 <- which(wgts_90cm[,2] < sp_w_inf)

bp90cm_histsoc <- tibble(time_steps, ymd_steps, tcb(projection_steady_F, min_w, max_w, sp90))
colnames(bp90cm_histsoc) <- c("Months.Since.1Jan1840", "Year.Month.Day", "bp90cm.gm-2")
bp90cm_nat <- tibble(time_steps, ymd_steps, tcb(projection_steady_noF, min_w, max_w, sp90))
colnames(bp90cm_nat) <- c("Months.Since.1Jan1840", "Year.Month.Day", "bp90cm.gm-2")

cp90cm_histsoc <- tibble(time_steps, ymd_steps, tc(projection_steady_F, sp90))
colnames(cp90cm_histsoc) <- c("Months.Since.1Jan1840", "Year.Month.Day", "cp90cm.gm-2")
```

Plot the output to make sure it looks reasonable in comparison to previous FishMIP runs, plotting each 
scenario separately because they're of such different orders of magnitude:
```{r}
#tcb - confused by order of magnitude greater than last GFDL-forced scenario, episodic nature of output.
par(mfrow = c(2,1))
plot(ymd_steps, tcb_histsoc$`tcb.gm-2`, type = "l")
plot(ymd_steps, tcb_nat$`tcb.gm-2`, type = "l")
```

```{r}
# tbp - confused by order of magnitude greater than last GFDL-forced scenario, episodic nature of output.
par(mfrow = c(2,1))
plot(ymd_steps, tpb_histsoc$`tpb.gm-2`, type = "l")
plot(ymd_steps, tpb_nat$`tpb.gm-2`, type = "l")
```

```{r}
# tcblog10 - order of magnitude greater than last GFDL-forced scenario (except where noted otherwise)
par(mfrow = c(2,1))
plot(ymd_steps, tcblog10_histsoc$`1to10g.gm-2`, type = "l")
plot(ymd_steps, tcblog10_nat$`1to10g.gm-2`, type = "l")

par(mfrow = c(2,1))
plot(ymd_steps, tcblog10_histsoc$`10to100g.gm-2`, type = "l")
plot(ymd_steps, tcblog10_nat$`10to100g.gm-2`, type = "l")

par(mfrow = c(2,1))
plot(ymd_steps, tcblog10_histsoc$`100gto1kg.gm-2`, type = "l")
plot(ymd_steps, tcblog10_nat$`100gto1kg.gm-2`, type = "l")

par(mfrow = c(2,1))
plot(ymd_steps, tcblog10_histsoc$`1to10kg.gm-2`, type = "l") 
plot(ymd_steps, tcblog10_nat$`1to10kg.gm-2`, type = "l") # extremely similar

par(mfrow = c(2,1))
plot(ymd_steps, tcblog10_histsoc$`10to100kg.gm-2`, type = "l") 
plot(ymd_steps, tcblog10_nat$`10to100kg.gm-2`, type = "l") 

par(mfrow = c(2,1))
plot(ymd_steps, tcblog10_histsoc$`GT100kg.gm-2`, type = "l") 
plot(ymd_steps, tcblog10_nat$`GT100kg.gm-2`, type = "l") 
```

```{r}
# tc - order of magnitude greater and episodic
par(mfrow = c(2,1))
plot(ymd_steps, tc_histsoc$`tc.gm-2`, type = "l") 
```

```{r}
# tic - not requested in prior rounds
par(mfrow = c(2,1))
plot(ymd_steps, tic_histsoc$`tic.gm-2`, type = "l") 
```

```{r}
# tclog10 
# no catch 1 - 100g as before
# 100g and greater: order of magnitude greater
par(mfrow = c(2,1))
plot(ymd_steps, tclog10_histsoc$`1to10g.gm-2`, type = "l")

par(mfrow = c(2,1))
plot(ymd_steps, tclog10_histsoc$`10to100g.gm-2`, type = "l")

par(mfrow = c(2,1))
plot(ymd_steps, tclog10_histsoc$`100gto1kg.gm-2`, type = "l")

par(mfrow = c(2,1))
plot(ymd_steps, tclog10_histsoc$`1to10kg.gm-2`, type = "l")

par(mfrow = c(2,1))
plot(ymd_steps, tclog10_histsoc$`10to100kg.gm-2`, type = "l")

par(mfrow = c(2,1))
plot(ymd_steps, tclog10_histsoc$`GT100kg.gm-2`, type = "l")
```

```{r}
# bp30to90cm - order of magnitude greater, though nat more similar
par(mfrow = c(2,1))
plot(ymd_steps, bp30to90cm_histsoc$`bp30to90cm.gm-2`, type = "l")
plot(ymd_steps, bp30to90cm_nat$`bp30to90cm.gm-2`, type = "l")
```

```{r}
# bp90cm - order of magnitude greater
par(mfrow = c(2,1))
plot(ymd_steps, bp90cm_histsoc$`bp90cm.gm-2`, type = "l")
plot(ymd_steps, bp90cm_nat$`bp90cm.gm-2`, type = "l")
```

```{r}
# cp30to90cm - order of magnitude greater
par(mfrow = c(2,1))
plot(ymd_steps, cp30to90cm_histsoc$`cp30to90cm.gm-2`, type = "l")
```

```{r}
# cp90cm - order of magnitude greater
par(mfrow = c(2,1))
plot(ymd_steps, cp90cm_histsoc$`cp90cm.gm-2`, type = "l")
```

Create some handy file name strings, for sanity:
```{r}
pre_var_histsoc <- c("mizer_gfdl-mom6-cobalt2_none_obsclim_histsoc_15arcmin_")
pre_var_nat <- c("mizer_gfdl-mom6-cobalt2_none_obsclim_nat_15arcmin_")
post_var <- c("_hawaii_monthly_1961_2010.csv")
```

Save each variable to its own file:
```{r}
write_csv(tcb_histsoc, file = paste("Output/", pre_var_histsoc, "tcb", post_var, sep = ""))
write_csv(tcb_nat, file = paste("Output/", pre_var_nat, "tcb", post_var, sep = ""))

write_csv(tcblog10_histsoc, file = paste("Output/", pre_var_histsoc, "tcblog10", post_var, sep = ""))
write_csv(tcblog10_nat, file = paste("Output/", pre_var_nat, "tcblog10", post_var, sep = ""))

write_csv(tpb_histsoc, file = paste("Output/", pre_var_histsoc, "tpb", post_var, sep = ""))
write_csv(tpb_nat, file = paste("Output/", pre_var_nat, "tpb", post_var, sep = ""))

write_csv(tc_histsoc, file = paste("Output/", pre_var_histsoc, "tc", post_var, sep = ""))

write_csv(tic_histsoc, file = paste("Output/", pre_var_histsoc, "tic", post_var, sep = ""))

write_csv(tclog10_histsoc, file = paste("Output/", pre_var_histsoc, "tclog10", post_var, sep = ""))

write_csv(bp30to90cm_histsoc, file = paste("Output/", pre_var_histsoc, "bp30to90cm", post_var, sep = ""))
write_csv(bp30to90cm_nat, file = paste("Output/", pre_var_nat, "bp30to90cm", post_var, sep = ""))

write_csv(bp90cm_histsoc, file = paste("Output/", pre_var_histsoc, "bp90cm", post_var, sep = ""))
write_csv(bp90cm_nat, file = paste("Output/", pre_var_nat, "bp90cm", post_var, sep = ""))

write_csv(cp30to90cm_histsoc, file = paste("Output/", pre_var_histsoc, "cp30to90cm", post_var, sep = ""))

write_csv(cp90cm_histsoc, file = paste("Output/", pre_var_histsoc, "cp90cm", post_var, sep = ""))
```
